BIP Architectures and Open pNets

### 2019-02-01 2018-Arch-pNets/arch-pNets.tex
formula ?
Shorter pnets and ot

\paragraph{pNets definition.}
\label{section:pNets}
A pLTS is a labelled transition system with variables that can be
used to characterise a state, as part of actions, guards, and
assignments. 
%Without loss of generality and to simplify the formalisation, we suppose 
%here that 
%variables are local to each 
%state: each state has its set of variables disjoint from the others. Transmitting 
%variable values from one state to the other can be done by explicit assignment. 
%Similarly, to simplify the management of variables and without loss of expressivity, we 
%suppose that transitions looping to the same state does not do assignments.
%Note that we make no assumption on the finiteness of the set of states, nor
%on the finite branching of the transition relation.
We need to introduce input variables to reason about the data flow between actions.
We  define the set of actions a pLTS can use ($a$
ranges over action labels):\\
% Action terms are  of the form 
$\alpha=a(?x_i^{i \in I}, 
\Expr_j^{j \in J})$, where $?x_i^{i \in I}$ are input variables, $\Expr_j^{j \in J}$ 
are expressions.

@@

\item $\symb{SV}_k^{k\in K}$ is a set of
  synchronisation vectors. % (\hl{$K\in\I_\P$}). 
$\forall k\!\in\! K,
  \symb{SV}_k\!=\SV{\!\alpha_{l}^{l\in I_k \uplus J_k}}{\alpha'_k}{g_k}$, where
  $\alpha'_k\in \actions{\variables}$, $I_k\subseteq I$, $J_k\subseteq J$, and 
  $\vars(\alpha'_k)\subseteq \bigcup_{l\in I_k\uplus 
  J_k}{\vars({\alpha_l})}$. The global action of a vector $\symb{SV}_k$ is
$\alpha'_k$. The Boolean expression $g_k $, such that $\vars(g_k)\subseteq \bigcup_{l\in 
I_k\uplus J_k}{\vars({\alpha_l})}$, is a guard associated to the vector.

\end{itemize}
%We define the set of holes and leaves of a pNet as follows:
%  \begin{itemize}
%\item
The set of \emph{holes} $\Holes(\pNet_{i_1})$ of a pNet is the indexes of the holes of the pNet 
itself plus the indexes of all the holes of its subnets (we suppose those indexes 
disjoints). A pNet $Q$ is \emph{closed} if it has no hole: $\Holes(Q)=\emptyset$; else it
is said to be \emph{open}.
%
%defined inductively; the sets of holes
%in a pNet node and its subnets are all disjoint:
%  \[\begin{array}{l}
%\Holes(\mylangle S,s_0, \to\myrangle) \!=\! \emptyset \\
%\Holes(\mylangle \pNet_i^{i\in I}\!,\Sort_j^{j\in J}\!, \overline{\symb{SV}}\myrangle) 
%=J\uplus{\displaystyle \bigcup_{i\in 
%I}\Holes(\pNet_i)}\\
%\forall i\in I.\, \Holes(\pNet_i)\cap J=\emptyset\\
%\forall i_1,i_2\in I.\,i_1\neq i_2\Rightarrow  
%\Holes(\pNet_{i_1})\cap\Holes(\pNet_{i_2})=\emptyset
%\end{array}\]
%\item
The set of \emph{leaves} of a pNet is the set of all pLTSs occurring in the structure, as an 
indexed family of the form $\Leaves(\pNet)= \mylangle \pNet_i \myrangle^{i \in L}$.
%\end{itemize}

@@

\paragraph{Operational Semantics for Open pNets.}
\label{section:op-semantics}

\noteLH{a pred is a g a post is a u (modified macros!)}
\renewcommand{\Pred}{g}
\renewcommand{\Post}{u}
The semantics of an open pNets is expressed as an open automaton, i.e., an automaton where each transition composes the actions of several holes, 
the transition occurs if some predicates hold, and can involve a 
 state modifications.
%\TODO{adopt a uniform notation for open transitions, almost each instance has a 
%different 
%notation! I suggest p,l,pr,po using \{\} for p,l,po as they are sets}
\begin{definition}[Open transition]
	\label{defn:OpenTransitions}
	An \emph{open transition} over a
	set $J$ of holes  and a set of states $\mathcal{S}$ is 
	a structure of the form:
\\[-2ex]	
	\begin{mathpar}
	\inferrule*[myfraction=\reddottedrule]
	{	\beta_j^{j\in J}, \Pred, \Post}
	{s \OTarrow {{\alpha}}s'}
	\end{mathpar}
	Where $s, s'\in\mathcal{S}$ and $\beta_j\in\actions{\variables}$
        is an action of the hole $j$. $v$ is a variable denoting the resulting action
        of this open transition. \Pred\ is a predicate 
	over the different variables of the
	terms, labels, and states  $\beta_j$, $s$, $\alpha$. $\Post\in 
	\assigns{\variables}$\ is 
	a set of assignments that are the effects of the transition.
Open transitions are identified modulo logical equivalence on their predicate.
\end{definition}
%\begin{definition}[Open transitions]
%	\label{def:OpenTransitions}
%	An \emph{open transition} over a set \noteSB{In the architecture section, I say that 
%we omit indices on $\goesto{}$ and that they are always clear from the 
%context.}\hl{$(S_i,s_{0 i}, \rightarrow_i)^{i\in
%	I}$} of LTSs, a
%	set $J$ of holes with sorts $\Sort_j^{j\in J}$, and a set of states $\mathcal{S}$ is 
%	a structure of the form:	
%	\begin{mathpar}
%	\inferrule*[myfraction=\reddottedrule]
%	{\{s_i~{\xrightarrow{a_i}}_i ~s_i^{\prime}\}^{i\in I},
%		\{\xrightarrow{b_j}_j\}^{j\in J}, \Pred, \Post}
%	{s \OTarrow {v}s'}
%	\end{mathpar}
%	Where $s, s'\in\mathcal{S}$ and for all
%        $i\in I$, $s_i{\xrightarrow{a_i}}_i s_i^{\prime}$ is a transition of the
%	LTS $(S_i,s_{0 i}, \rightarrow_i)$, and \noteSB{Why insist on ``transition'' and 
%$\goesto{b_j}_j$? Why not simply ``$b_j$ is an action in the sort 
%$\Sort_j$?}\hl{$\xrightarrow{b_j}_j$
%        is a transition of the hole $j$}, for any action $b_j$ in the
%        sort $\Sort_j$.  
%        of this open transition. \Pred\ is a predicate 
%	over the different variables of the
%	terms, labels, and states $s_i$, $b_j$, $s$, $v$. \Post\ is a set of equations that 
%	hold \emph{after the open transition}, they are represented as a substitution of the 
%	form $\{x_k\gets \Expr_k\}^{k\in K}$ \noteLH{considering the level of abstraction I 
%	would like to just put $\Post_=e$}
%	where $x_k$ are variables of $s'$, $s'_i$, and $e_k$ are expressions over the other 
%	variables of the open transition.
%\end{definition}

@@

\begin{example}[An open-transition]
  \label{Example:TimerOT3}
  The \texttt{Failure TimerMonitor} pNet of \Fig{FailureTimer:pNet} has \replaceSB{2 controllers 
  and 2 holes. O}{two coordinators and four dangling ports, which can be grouped into two holes (see \secn{encoding} below).  Disregarding the maximal progress priority, o}ne of its possible open-transitions is:\\[-2ex] \noteLH{maximal progress undefined}
%%   \footnote{Note the underlined action notation that we use to represent synchronised actions, i.e. actions that are the result of a synchronisation and that we declare as \emph{non-observable}. Those actions are only used for monitoring purposes and statement of properties but cannot be synchronised again in a synchronisation vector. This allows us to define a weak bisimilarity equivalence with p=compositional properties.}
%%   : 
%% \noteLH{underline notation might be a bit too much for this paper : see at the end if we remove it or not (will we use it?)}
% \smallskip
%$  OT_1  = \openrule
%{\emptyset, \{t:=0, z:=TopInt, zone:=[minZone,maxZone]\} }
%{\ostate{00} \OTarrow{\underline{\alpha_4}} \ostate{11}}
%$
%\medskip
\begin{small}\[ \openrule
  {\{B\mapsto fail(b0)\},\\
    t<GetMax(z) \land b1=true \land b1=>b0 \land z'= (z \cap (b1?zone;TopInt)+t) {\color{blue}\land (t' = t) \land (zone' = zone)},\\
    \{z:=z', {\color{blue}t := t'+1, zone := zone'}\} }
  {\ostate{11} \OTarrow{\alpha_0} \ostate{12}}
  \]
\end{small}\vspace{-1.5ex}
%
%The first one, $OT_1$, corresponds to the simultaneous initialisation
%of both controlers. No hole is active at that time, and the resulting
%global action is denoted $\alpha_4$.

It corresponds to the firing of the \replaceSB{connector
$''tick, fail, fail''$}{interaction $\listset{\NameTimer.\PortTick, \PortFail, \NameCtrl.\PortFail}$}. It states that, if the hole $B$ performs an action $fail$, and the (complex) guard is satisfied, then the variable $z$ is set to $z1$ (as specified by the guard), then the transition below the dotted line occurs and a global action labelled $\alpha_0$ occurs. Only the second pLTS performs a transition from state $1$ to state $2$.
The guard corresponds both to the guard of this transition and to synchronisation vector constraints.

>>>>

%\paragraph{Definitions.}
%\label{section:pNets}
A pLTS is a labelled transition system with variables that can be
used to characterise a state, as part of actions, guards, and
assignments. 
%Without loss of generality and to simplify the formalisation, we suppose 
%here that 
%variables are local to each 
%state: each state has its set of variables disjoint from the others. Transmitting 
%variable values from one state to the other can be done by explicit assignment. 
%Similarly, to simplify the management of variables and without loss of expressivity, we 
%suppose that transitions looping to the same state does not do assignments.
%Note that we make no assumption on the finiteness of the set of states, nor
%on the finite branching of the transition relation.
We need to introduce input variables to reason about the data flow between actions.
We  define the set of actions a pLTS can use ($a$
ranges over action labels):\\
% Action terms are  of the form 
$\alpha=a(?x_i^{i \in I}, 
\Expr_j^{j \in J})$, where $?x_i^{i \in I}$ are input variables, $\Expr_j^{j \in J}$ 
are expressions.

@@

\item $\symb{SV}_k^{k\in K}$ is a set of
  synchronisation vectors. % (\hl{$K\in\I_\P$}). 
$\forall k\!\in\! K,
  \symb{SV}_k\!=\SV{\!\alpha_{l}^{l\in I_k \uplus J_k}}{\alpha'_k}{g_k}$, where
  $\alpha'_k\in \actions{\variables}$, $I_k\subseteq I$, $J_k\subseteq J$, and 
  $\vars(\alpha'_k)\subseteq \bigcup_{l\in I_k\uplus 
  J_k}{\vars({\alpha_l})}$. The global action  is
$\alpha'_k$,  $g_k$ is a guard associated to the vector.

\end{itemize}
%We define the set of holes and leaves of a pNet as follows:
%  \begin{itemize}
%\item
The set of \emph{holes} $\Holes(\pNet)$ of a pNet is the indexes of the holes of the pNet 
itself plus the indexes of all the holes of its subnets (we suppose those indexes 
disjoints). A pNet $Q$ is \emph{closed} if it has no hole: $\Holes(Q)=\emptyset$; else it
is said to be \emph{open}.
%
%defined inductively; the sets of holes
%in a pNet node and its subnets are all disjoint:
%  \[\begin{array}{l}
%\Holes(\mylangle S,s_0, \to\myrangle) \!=\! \emptyset \\
%\Holes(\mylangle \pNet_i^{i\in I}\!,\Sort_j^{j\in J}\!, \overline{\symb{SV}}\myrangle) 
%=J\uplus{\displaystyle \bigcup_{i\in 
%I}\Holes(\pNet_i)}\\
%\forall i\in I.\, \Holes(\pNet_i)\cap J=\emptyset\\
%\forall i_1,i_2\in I.\,i_1\neq i_2\Rightarrow  
%\Holes(\pNet_{i_1})\cap\Holes(\pNet_{i_2})=\emptyset
%\end{array}\]
%\item
The set of \emph{leaves} of a pNet is the set of all pLTSs occurring in the structure, as an 
indexed family of the form $\Leaves(\pNet)= \mylangle \pNet_i \myrangle^{i \in L}$.
%\end{itemize}

@@

%\paragraph{Operational Semantics.}
%\label{section:op-semantics}
\renewcommand{\Pred}{g}
\renewcommand{\Post}{u}
The semantics of an open pNets is expressed as  an automaton where each transition composes the actions of several holes, 
the transition occurs if some predicates hold, and can involve  
 state modifications.
%\TODO{adopt a uniform notation for open transitions, almost each instance has a 
%different 
%notation! I suggest p,l,pr,po using \{\} for p,l,po as they are sets}
\begin{definition}[Open transition]
	\label{defn:OpenTransitions}
	An \emph{open transition} over a
	set $J$ of holes  and a set of states $\mathcal{S}$ is 
	a structure of the form:
\\[-2ex]	
	\begin{mathpar}
	\inferrule*[myfraction=\reddottedrule]
	{	\beta_j^{j\in J}, \Pred, \Post}
	{s \OTarrow {{\alpha}}s'}
	\end{mathpar}
	Where $s, s'\in\mathcal{S}$ and $\beta_j\in\actions{\variables}$
        is an action of the hole $j$. $v$ is a variable denoting the resulting action
        of this open transition. \Pred\ is a predicate 
	over the different variables of the
	terms, labels, and states  $\beta_j$, $s$, $\alpha$. $\Post\in 
	\assigns{\variables}$\ is 
	a set of assignments that are the effects of the transition.
Open transitions are identified modulo logical equivalence on their predicate.
\end{definition}
%\begin{definition}[Open transitions]
%	\label{def:OpenTransitions}
%	An \emph{open transition} over a set \noteSB{In the architecture section, I say that 
%we omit indices on $\goesto{}$ and that they are always clear from the 
%context.}\hl{$(S_i,s_{0 i}, \rightarrow_i)^{i\in
%	I}$} of LTSs, a
%	set $J$ of holes with sorts $\Sort_j^{j\in J}$, and a set of states $\mathcal{S}$ is 
%	a structure of the form:	
%	\begin{mathpar}
%	\inferrule*[myfraction=\reddottedrule]
%	{\{s_i~{\xrightarrow{a_i}}_i ~s_i^{\prime}\}^{i\in I},
%		\{\xrightarrow{b_j}_j\}^{j\in J}, \Pred, \Post}
%	{s \OTarrow {v}s'}
%	\end{mathpar}
%	Where $s, s'\in\mathcal{S}$ and for all
%        $i\in I$, $s_i{\xrightarrow{a_i}}_i s_i^{\prime}$ is a transition of the
%	LTS $(S_i,s_{0 i}, \rightarrow_i)$, and \noteSB{Why insist on ``transition'' and 
%$\goesto{b_j}_j$? Why not simply ``$b_j$ is an action in the sort 
%$\Sort_j$?}\hl{$\xrightarrow{b_j}_j$
%        is a transition of the hole $j$}, for any action $b_j$ in the
%        sort $\Sort_j$.  
%        of this open transition. \Pred\ is a predicate 
%	over the different variables of the
%	terms, labels, and states $s_i$, $b_j$, $s$, $v$. \Post\ is a set of equations that 
%	hold \emph{after the open transition}, they are represented as a substitution of the 
%	form $\{x_k\gets \Expr_k\}^{k\in K}$ \noteLH{considering the level of abstraction I 
%	would like to just put $\Post_=e$}
%	where $x_k$ are variables of $s'$, $s'_i$, and $e_k$ are expressions over the other 
%	variables of the open transition.
%\end{definition}

@@

\begin{example}[An open-transition]
  \label{Example:TimerOT3}
We now briefly illustrate an open transition:
\begin{small}\[ \openrule
  {\{B\mapsto fail(b)\},
    t<Max \land b=true ,
    \{t' := t+1\} }
  {\ostate{11} \OTarrow{\alpha(t)} \ostate{12}}\vspace{-1.5ex}
  \]
\end{small}
%The first one, $OT_1$, corresponds to the simultaneous initialisation
%of both controlers. No hole is active at that time, and the resulting
%global action is denoted $\alpha_4$.
It requires the hole at label $B$ to fire a $fail(b)$ action, with the condition $b=true$ and the state variable $t<Max$. In this case, the global pNet goes from state $\ostate{11}$ to state $\ostate{12}$ where $t'=t+1$. This emits a global action $\alpha(t)$.
%It corresponds to the firing of the \replaceSB{connector
%$''tick, fail, fail''$}{interaction $\listset{\NameTimer.\PortTick, \PortFail, \NameCtrl.\PortFail}$}. It states that, if the hole $B$ performs an action $fail$, and the (complex) guard is satisfied, then the variable $z$ is set to $z1$ (as specified by the guard), then the transition below the dotted line occurs and a global action labelled $\alpha_0$ occurs. Only the second pLTS performs a transition from state $1$ to state $2$.
%The guard corresponds both to the guard of this transition and to synchronisation vector constraints.
